Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
almost done up until results sections

The final principle when discussing parallel performance is the potential existence of an Amdahl speedup limit. Amdahl's Law seeks to describe the maximum parallel speedup in an overall piece of code when the problem is not embarassingly parallel --- not all of the system can be divided up in parallel. The parallel in time methods must always have some serial component due to the time-dependent nature of the ODE's which will fall under the limitations stated by Amdahl's Law.  His law expresses both best time given $n$ processors as well as the speedup for $n$ processors as a function of $B$, the fractional amount of the algorithm that can be parallelized:

@@

We look to compute a theoretical maximum speedup enabled by the nature of the Parareal algorithm.  As this is a tradeoff between error (with parameter k being the number of iterations of corrections, each which takes another cycle of $g_{\Delta t}$, we seek to express the tradeoff as a function of the quality factor, $Q$, the number of iterations, $k$ and the number of processors $n$.  Our quality factor is defined as the multiplicative factor of additional time necessary to TODO

Then, the runtime of Parareal is, assuming negligible overhead and ignoring communication speeds and letting $t$ be the time in seconds to compute the coarse $g_{\Delta t}$:

@@

Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound (with $B$ of 1):

\[
est
\]

@@

@@

We wished to explore certain tests that would span the space of possible
parameters to the parareal as well as possible applications in different ODE
systems. Because there were so many different versions and parameters for the Parareal to test,

\subsection{Expectations}
TODO WESLEY

>>>>

The final principle when discussing parallel performance is the potential existence of an Amdahl speedup limit. Amdahl's Law seeks to describe the maximum parallel speedup in an overall piece of code when the problem is not embarassingly parallel --- not all of the system can be divided up in parallel. The parallel in time methods must always have some serial component due to the time-dependent nature of the ODE's which will fall under the limitations stated by Amdahl's Law.  His law expresses both best time given $n$ processors as well as the speedup for $n$ processors as a function of $B$, the fractional amount of the algorithm that cannot be parallelized---due to strict serial constraints:

@@

We look to compute a theoretical maximum speedup enabled by the nature of the Parareal algorithm.  As this is a tradeoff between error (with parameter k being the number of iterations of corrections, each which takes another cycle of $g_{\Delta t}$, we seek to express the tradeoff as a function of the quality factor, $Q$, the number of iterations, $k$ and the number of processors $n$.  Our quality factor is defined as the multiplicative factor of additional time necessary to compute the solution using the $g_{\textnormal{fine}}$ scheme instead of the $g_{\Delta t}$ method.

Then, the runtime of Parareal is, assuming negligible overhead and ignoring communication bandwidth and letting $t$ be the time in seconds to compute the coarse $g_{\Delta t}$:

@@

Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of (with $B$ of 0):

\[
S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
\]

With embarassingly parallel problem setups, we can optimally achieve up to $n$ times speedup with $n$ processors though this number if quite ideal and will be very difficult to approach (large scaling will have different memory management and greater communication overheads and costs).

@@

We wished to explore certain tests that would demonstrate the effects of changing various parameters be it from the solution schemes in the parareal to the exact system we were trying to solve.

Due to the sheer number of things to explore, we usually try to compare each run to a benchmark we decided was most useful for comparison's sake and the list of parameters we wished to test are:

\begin{itemize}
\item
\item
\end{itemize}

@@

\subsection{Expectations}
We hoped to show that under the properly tuned parameters (namely the $k$ number of iterations),
